Non-Vertical Cultural Transmission, Assortment, \\and the Evolution of Cooperation

### 2020-11-11 ms/main.tex
formatting

The stable population mean fitness is $\bar{w}^* = 1 + \hat{p}^*(b-c)$, which is increasing in $\hat{p}^*$. 
In turn, $\hat{p}^*$ is also increasing in $\alpha$ (Appendix~\ref{sec:appendixC}).
Therefore, if social association increases, the population mean fitness also increases.
But can an allele that increases social association evolve? 
To answer this question, we extend our model to include a ``modifier locus''~\citep{Liberman1986,Liberman1988} that determines social association, but has no direct effect on fitness.
Assume that the modifier locus has two alleles, $M$ and $m$, which induce social associations $\alpha_1$ and $\alpha_2$, respectively.
Suppose that the population evolves to a stable equilibrium when only allele $M$ is present.
We check the local stability of this equilibrium to the invasion of the modifier allele $m$; this stablity is called ``external stability'', see \citet{Altenberg2017}.

The full analysis is given in Appendix~\ref{sec:appendixD}, and following result summarizes the outcomes of such invasions.
\\

@@

Note that this reduction principle entails that successful invasions will reduce the frequency of cooperation, as well as the population mean fitness (\autoref{fig:invasion}).
Furthermore, if we assume that modifier alleles with decreased social association appear and invade the population from time to time, then, social association will continue to decrease, further reducing the frequency of cooperation and the population mean fitness. This evolution will proceed as long as stable polymorphism occurs, that is, as long as $a_2 < \alpha < a_1$ (Remark~\ref{remark:intermediate_association_res3}
, \autoref{fig:result2}c). 
Thus, we can expect social association to eventually drop to $a_2$, the frequency of cooperation to drop to zero, and the population mean fitness to drop to one.
\\

>>>>

The stable population mean fitness is $\bar{w}^* = 1 + \hat{p}^*(b-c)$, which is increasing in $\hat{p}^*$. 
In turn, $\hat{p}^*$ is also increasing in $\alpha$ (Appendix~\ref{sec:appendixC}).
Therefore, if social association increases, the population mean fitness also increases.
But can an allele that increases social association evolve?

To answer this question, we extend our model to include a ``modifier locus''~\citep{Liberman1986,Liberman1988} that determines social association, but has no direct effect on fitness.
We assume that the modifier locus has two alleles, $M$ and $m$, which induce social associations $\alpha_1$ and $\alpha_2$, respectively.
Suppose that the population evolves to a stable equilibrium when only allele $M$ is present.
We check the local stability of this equilibrium to the invasion of the modifier allele $m$; this stablity is called ``external stability'', see \citet{Altenberg2017}.
The full analysis is given in Appendix~\ref{sec:appendixD}, and the following result summarizes the outcomes of such invasions.
\\

@@

Note that this reduction principle entails that successful invasions will reduce the frequency of cooperation, as well as the population mean fitness (\autoref{fig:invasion}).
Furthermore, if we assume that modifier alleles with decreased social association appear and invade the population from time to time, then, the social association will continue to decrease, further reducing the frequency of cooperation and the population mean fitness. This evolution will proceed as long as stable polymorphism occurs, that is, as long as $a_2 < \alpha < a_1$ (Remark~\ref{remark:intermediate_association_res3}, \autoref{fig:result2}c). 
Thus, we can expect social association to eventually drop to $a_2$, the frequency of cooperation to drop to zero, and the population mean fitness to drop to one.
